humanCdGF_predict_cfu

### 2016-11-23 code/CdGF_AUCRF.Rmd
using full shared files

meta_file   <- '~/Documents/Github/Schubert_humanCdGF_XXXX_2016/data/process/human_CdGF_metadata.txt'
shared_file <- '~/Documents/Github/Schubert_humanCdGF_XXXX_2016/data/process/human_CdGF.an.unique_list.0.03.subsample.shared'
tax_file <- '~/Documents/Github/Schubert_humanCdGF_XXXX_2016/data/process/gf_new.an.taxonomy'
tax_function <- '~/Documents/Github/Schubert_humanCdGF_XXXX_2016/code/tax_level.R'

@@

# subset shared
#Create df with relative abundances
stool_shared <- shared_file[rownames(meta_file[meta_file$day %in% 0, ]), ]
# use %in% instead of == in order to ignore NA
# rename day 0 community rows by mouse id - since meta rows have day in row names, use mouse id for both
rownames(stool_shared) <- meta_file$mouse_id[rownames(meta_file) %in% 
                                               rownames(stool_shared) ]
rel_abund_stool <- 100 * stool_shared / unique(apply(stool_shared, 1, sum))
#Create vector of OTUs with median abundances >1%
otu_presence <- ifelse(stool_shared > 0, T, F) 
OTUs_present <- colnames(stool_shared[ , apply(otu_presence, 2, sum) > round(otu_presence_cutoff*nrow(stool_shared))])
#df of OTUs with present in >10%
rel_abund_stool <- rel_abund_stool[ ,OTUs_present]

>>>>

meta_file   <- '~/Documents/Github/Schubert_humanCdGF_XXXX_2016/data/process/human_CdGF_metadata.txt'
shared_file <- '~/Documents/Github/Schubert_humanCdGF_XXXX_2016/data/process/humanCdGF_full.shared'
tax_file <- '~/Documents/Github/Schubert_humanCdGF_XXXX_2016/data/process/gf_new.an.taxonomy'
tax_function <- '~/Documents/Github/Schubert_humanCdGF_XXXX_2016/code/tax_level.R'

@@

# subset shared
#Create df with relative abundances
stool_shared <- shared_file[rownames(meta_file[meta_file$day %in% 0, ]), ]
# use %in% instead of == in order to ignore NA
# rename day 0 community rows by mouse id - since meta rows have day in row names, use mouse id for both
rownames(stool_shared) <- meta_file$mouse_id[rownames(meta_file) %in% 
                                               rownames(stool_shared) ]
rel_abund_stool <- 100 * stool_shared / apply(stool_shared, 1, sum)
#Create vector of OTUs with median abundances >1%
otu_presence <- ifelse(rel_abund_stool >= (1/2000), T, F) 
OTUs_present <- colnames(rel_abund_stool[ , apply(otu_presence, 2, sum) > round(otu_presence_cutoff*nrow(rel_abund_stool))])
#df of OTUs with present in >10%
rel_abund_stool <- rel_abund_stool[ ,OTUs_present]

### 2016-11-23 code/HumanCdGF_predict_cfu.Rmd
using full shared files

```{r setup data}
# read in files
setwd('~/Documents/Github/Schubert_humanCdGF_XXXX_2016/')
meta_data   <- 'data/process/human_CdGF_metadata.txt'
shared_file <- 'data/process/human_CdGF.an.unique_list.0.03.subsample.shared'
tax_file <- 'data/process/gf_new.an.taxonomy'
tax_function <- '~/Documents/Github/Schubert_humanCdGF_XXXX_2016/code/tax_level.R'

@@

# set parameters
seed <- 1
n_trees <- 1001
n_otus <- 12
iters <- 100
cor_cutoff <- 0.75

@@

# create shared with relative abundances over 1%
mean_otu_counts <- unique(apply(shared_file, 1, sum))
rel_abund_shared <- shared_file
rel_abund_shared_01 <- rel_abund_shared[ , apply(rel_abund_shared, 2, max) > 1]
# remove correlative otus
rel_abund_cor <- cor(rel_abund_shared_01, method = 'spearman')
rownames(rel_abund_cor) <- colnames(rel_abund_shared_01) # add row names to identify otus correlated with features
cor_otus <- findCorrelation(rel_abund_cor, cutoff = cor_cutoff)
rel_abund_01_filtr <- rel_abund_shared_01[,-cor_otus]
community_d0.df <- merge(shared_list_days, rel_abund_01_filtr, 
                         by.x = 'sample_id', by.y = 'row.names') %>% 
                    select(-sample_id)
```

@@

rfe_ctrl <- rfeControl(functions = rfRFE,
                   method = "repeatedcv",
                   repeats = 10,
                   verbose = FALSE)
subsets <- c(1:20, c(seq(25, 85, 10)))
set.seed(seed)
rf_cfu_model <- rfe(cfu_shared_colonized.df[ , -1], cfu_shared_colonized.df$log_cfu_med,
                 sizes = subsets,
                 rfeControl = rfe_ctrl)
top_important_OTU <- predictors(rf_cfu_model)

@@

layout(matrix(c(1:9), ncol=3, byrow=TRUE))
for(otu in seq_along(importance_om$tax)){
    partialPlot(rf_n_features, data.frame(cfu_shared_colonized.df), x.var=rownames(importance_om)[otu],
                xlab='', main = NULL)
    if(length(features_cor[[otu]]) > 1){
      title(main = paste0('Unidentifiable Feature\n(', 
                          get_tax(row_list = rownames(importance_om)[otu])$tax_label,
                          ' Shown)'))
      mtext(side = 1, line = 3, cex = 0.6,
        paste('Correlated features: \n', 
              paste(get_tax(row_list = features_cor[[otu]])$tax_label, collapse = ', ')))
      } else {
      title(paste(get_tax(row_list = rownames(importance_om)[otu])$tax_label))
        }
    par(new = T)
    plot(y = partial_df[ , 'log_cfu_med'],
         x = partial_df[ , rownames(importance_om)[otu]], 
         pch = NA, bg = 'gray', xlab = NA, ylab = NA, axes = F)
    axis(side = 4)
    mtext(side = 4, line = 3, 'Log10 CFU')
# To color points by mouse - Change pch to 21 and un-comment following lines
  lapply(seq_along(unique(partial_df$cage_id)), function(m){
    mouse <- unique(partial_df$cage_id)[m]
    points(y = partial_df[partial_df$cage_id %in% mouse, 'log_cfu_med'],
           x = partial_df[partial_df$cage_id %in% mouse , rownames(importance_om)[otu]], 
           pch = 21, cex = 2, bg = color[m], xlab = NA, ylab = NA, axes = F)
  })
}
```

@@

layout(matrix(c(1:9), ncol=3, byrow=TRUE))

cage_color <- rainbow(length(unique(cr_prediction$cage_id)))
for (otu in seq_along(cr_features)) {
    partialPlot(rf_col_rate, shared_cr_df, cr_features[otu],
                xlab='', main = NULL)
    if(length(cr_features_cor[[otu]]) > 1){
      title(main = paste0('Unidentifiable Feature\n(', 
                          get_tax(row_list = cr_features[otu])$tax_label,
                          ' Shown)'))
      mtext(side = 1, line = 3, cex = 0.6,
        paste('Correlated features: \n', 
              paste(get_tax(row_list = cr_features_cor[[otu]])$tax_label, collapse = ', ')))
      } else {
      title(paste(get_tax(row_list = cr_features[otu])$tax_label))
        }
  par(new = T)
    plot(y = shared_col_rate_df[ , 'ratio'],
         x = shared_col_rate_df[ , cr_features[otu]], 
         pch = NA, bg = 'gray', xlab = NA, ylab = NA, axes = F)
    axis(side = 4)
    mtext(side = 4, line = 3, 'Relative Colonization')
# To color points by mouse - Change pch to 21 and un-comment following lines
  lapply(seq_along(unique(shared_col_rate_df$cage_id)), function(m){
    mouse <- unique(shared_col_rate_df$cage_id)[m]
    points(y = shared_col_rate_df[shared_col_rate_df$cage_id %in% mouse, 'ratio'],
           x = shared_col_rate_df[shared_col_rate_df$cage_id %in% mouse , cr_features[otu]], 
           pch = 21, cex = 2, bg = color[m], xlab = NA, ylab = NA, axes = F)
  })
}
```

>>>>

```{r setup data}
# read in files
setwd('~/Documents/Github/Schubert_humanCdGF_XXXX_2016/')
meta_data   <- 'data/process/human_CdGF_metadata.txt'
shared_file <- 'data/process/humanCdGF_full.shared'
tax_file <- 'data/process/gf_new.an.taxonomy'
tax_function <- '~/Documents/Github/Schubert_humanCdGF_XXXX_2016/code/tax_level.R'

@@

# set parameters
seed <- 1
n_trees <- 1001
n_otus <- 12
iters <- 100
cor_cutoff <- 0.75
otu_presence_cutoff <- 0.1

@@

# create shared with relative abundances over 1%
rel_abund_shared <- 100 * shared_file / apply(shared_file, 1, sum)
otu_presence <- ifelse(rel_abund_shared >= (1/2000), T, F) 
rel_abund_shared_01 <- rel_abund_shared[ , apply(otu_presence, 2, sum) > round(otu_presence_cutoff*nrow(rel_abund_shared))]
# remove correlative otus
rel_abund_cor <- cor(rel_abund_shared_01, method = 'spearman')
rownames(rel_abund_cor) <- colnames(rel_abund_shared_01) # add row names to identify otus correlated with features
cor_otus <- findCorrelation(rel_abund_cor, cutoff = cor_cutoff)
rel_abund_01_filtr <- rel_abund_shared_01[,-cor_otus]
community_d0.df <- merge(shared_list_days, rel_abund_01_filtr, 
                         by.x = 'sample_id', by.y = 'row.names') %>% 
                    select(-sample_id)
```

@@

rfe_ctrl <- rfeControl(functions = rfRFE,
                   method = "repeatedcv",
                   repeats = 10,
                   verbose = FALSE)
subsets <- c(1:15, c(seq(20, 70, 10)))
set.seed(seed)
rf_cfu_model <- rfe(cfu_shared_colonized.df[ , -1], cfu_shared_colonized.df$log_cfu_med,
                 sizes = subsets,
                 rfeControl = rfe_ctrl)
top_important_OTU <- predictors(rf_cfu_model)

@@

layout(matrix(c(1:9), ncol=3, byrow=TRUE))
par(mar = c(7,4,4,2)+0.1)
for(otu in seq_along(importance_om$tax)){
    partialPlot(rf_n_features, data.frame(cfu_shared_colonized.df), x.var=rownames(importance_om)[otu],
                xlab='', main = NULL)
    if(length(features_cor[[otu]]) > 1){
      title(main = paste0('Unidentifiable Feature\n(', 
                          get_tax(row_list = rownames(importance_om)[otu])$tax_label,
                          ' Shown)'))
      mtext(side = 1, line = 6, cex = 0.6,
        paste('Correlated features: \n', 
              paste(paste0(get_tax(row_list = features_cor[[otu]])$tax_label, c(', ', ',\n ')), 
                    collapse = '')))
      } else {
      title(paste(get_tax(row_list = rownames(importance_om)[otu])$tax_label))
        }
    par(new = T)
    plot(y = partial_df[ , 'log_cfu_med'],
         x = partial_df[ , rownames(importance_om)[otu]], 
         pch = NA, bg = 'gray', xlab = NA, ylab = NA, axes = F)
    axis(side = 4)
    mtext(side = 4, line = 3, 'Log10 CFU')
# To color points by mouse - Change pch to 21 and un-comment following lines
  lapply(seq_along(unique(partial_df$cage_id)), function(m){
    mouse <- unique(partial_df$cage_id)[m]
    points(y = partial_df[partial_df$cage_id %in% mouse, 'log_cfu_med'],
           x = partial_df[partial_df$cage_id %in% mouse , rownames(importance_om)[otu]], 
           pch = 21, cex = 2, bg = color[m], xlab = NA, ylab = NA, axes = F)
  })
}
```

@@

layout(matrix(c(1:9), ncol=3, byrow=TRUE))
par(mar = c(7,4,4,2)+0.1)
cage_color <- rainbow(length(unique(cr_prediction$cage_id)))
for (otu in seq_along(cr_features)) {
    partialPlot(rf_col_rate, shared_cr_df, cr_features[otu],
                xlab='', main = NULL)
    if(length(cr_features_cor[[otu]]) > 1){
      title(main = paste0('Unidentifiable Feature\n(', 
                          get_tax(row_list = cr_features[otu])$tax_label,
                          ' Shown)'))
      mtext(side = 1, line = 6, cex = 0.6,
        paste('Correlated features: \n', 
              paste(paste0(get_tax(row_list = cr_features_cor[[otu]])$tax_label, c(', ', ',\n ')), 
                    collapse = '')))
      } else {
      title(paste(get_tax(row_list = cr_features[otu])$tax_label))
        }
  par(new = T)
    plot(y = shared_col_rate_df[ , 'ratio'],
         x = shared_col_rate_df[ , cr_features[otu]], 
         pch = NA, bg = 'gray', xlab = NA, ylab = NA, axes = F)
    axis(side = 4)
    mtext(side = 4, line = 3, 'Relative Colonization')
# To color points by mouse - Change pch to 21 and un-comment following lines
  lapply(seq_along(unique(shared_col_rate_df$cage_id)), function(m){
    mouse <- unique(shared_col_rate_df$cage_id)[m]
    points(y = shared_col_rate_df[shared_col_rate_df$cage_id %in% mouse, 'ratio'],
           x = shared_col_rate_df[shared_col_rate_df$cage_id %in% mouse , cr_features[otu]], 
           pch = 21, cex = 2, bg = color[m], xlab = NA, ylab = NA, axes = F)
  })
}
```
